---
title: "TEDS-A Descriptive Analysis, 2000-2021"
author: "Simone Wien "
date: "`r Sys.Date()`"
output: html_document

---

# Introduction 

The data set we will be looking at is the *Treatment Episode Data Set: Admissions (TEDS-A)* dataset, which is publicly available though the *Substance Abuse and Mental Health Services Administration (SAMSHA)*. TEDS-A contains admission-level substance use treatment information among people aged 12 years or older, and contains information on individual demographics and substance use characteristics. More information on TEDS-A can be found [here](https://www.datafiles.samhsa.gov/data-sources). 

The purpose of this analysis is to describe the sample of natally sexed females of reproductive age (15-49) by region, race, and insurance status in publicly funded drug treatments centers in the United States. We will be using data from 2000-2021, and the codebook can be found [here](https://www.datafiles.samhsa.gov/dataset/treatment-episode-data-set-admissions-2000-2021-teds-2000-2021-ds0001). 


```{r, set-chunk-opts, echo = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE #all options we'd like for the whole document
)
```

```{r packages}
#packages list 
pacman::p_load(data.table, #fast aggregation,large data
              tidyverse, #general data handling
              skimr, #summary statistics
              here, #easy file referencing
              ggplot2, #plots
              gtsummary,  #easy Table 1
              gridExtra, #arrange multiple ggplots
              hrbrthemes) #theme for histogram 
```

```{r, load-data}
here::i_am("code/report.Rmd")

data <- readRDS(
  file = here::here("data/data_clean.rds")
)
```

# Descriptive analysis

## Admission demographics

From 2000-2021, there were `r nrow(data)` admissions by women of reproductive age into publicly funded treatment centers across the United States. Below are select characteristics of the sample by Census region.

Approximately 8% of the total sample were pregnant at the time of admission, with noticeable differences by race and insurance status across regions.

```{r, table1}
table_one <- readRDS(
  file = here::here("tables/table_one.rds")
)
table_one
```


## Admission over time

Below is a figure showing the frequency of admissions over time by year, broken down by pregnancy status. At the national level, 1) admissions peak in 2008 and 2018, with a noticiable decline from 2018-2021, and 2) the proportion of pregnant, non-pregnant, missing pregnancy status individuals shift across time. 

```{r, figure1}
#| fig.align = "center",
#| out.width = "600px"
knitr::include_graphics(
  here::here("figures/figure_one.png")
)
```


